---
title: Analysing Serving Abilities at Grand Slams
author: Peter Tea
date: '2019-12-26'
categories:
  - Tennis
tags:
  - Bubble plot
  - Tennis
authors:
  - admin
output:
  html_document:
    keep_md: yes
image:
  caption: Photo by johnfo on Unsplash
  focal_point: Smart
summary: Visualizing the beautiful game of Tennis.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 2019 Serving Efficiency at Grand Slams

{{% alert note %}}
**Attribution:** Special thanks to [Jeff Sackmann](http://www.tennisabstract.com/) for scraping and providing the data freely on his [Github page](https://github.com/JeffSackmann/tennis_atp/blob/master/README.md).
{{% /alert %}}

### Recapping the 2019 Grand Slam season

With the conclusion of the 2019 Tennis season, the Big 3 (Nadal, Djokovic and Federer) all continued their domination during the largest stages of the season: the grand slam tournaments
  
  **Rafa Nadal** cruised to the *Roland Garros* chip winning the final match in four sets against Thiem, while also capturing the *US Open* championship in a 5-set thriller over Medvedev.
  
  Meanwhile, **Novak Djokovic** successfully defended his *Wimbledon* trophy against Roger Federer in a heavy-weight 5-set match, while also capturing the *Australian Open* chip against Nadal. 
  
  Finally, while **Roger Federer** did not win any grand slams in this 2019 season, he did reach the Wimbledon final as well as make some exciting runs at the other 3 grand slam tournaments. Not bad for the 38-year old Swiss! 


### How well did the Big 3 defend and break serves?

To analyse at a granular level how well the Big 3 performed during the 2019 grand slam tournaments, we look at their performances in **defending** their own service games and in **attacking** their opponent service games. A player who effectively defends their serve should win a high proportion of the games in which they serve first. Furthermore, a player who successfully attacks their opponents' serve should win a reasonable amount of break points against their opponent's serve.

To measure a players defensive and offensive ability, we look at the total number of service games defended and the total number of opponent serves broken respectively. Moreover, to account for opportunity, we divide these metrics by the total number of player service games and opponent service games respectively.


- Proportion of service games defended = # of service games won / Total # player of service games

- Proportion of opponent service games broken = # of service games broken / Total # of opponent service games

```{r gg, echo = FALSE, warning=FALSE,message=F}
#Proportion of service games defended = 
#$\frac{ \text{# of service games won}} {\text{Total # of service games}}$

#- Proportion of opponent service games broken = 
#\[\frac{ \text{# of service games broken}} {\text{Total # of opponent service games}} \]
```


### What do the numbers say?

```{r libraries, echo = FALSE, warning=FALSE,message=F}
library(dplyr)
library(ggplot2)
library(ggrepel)
library(plotly)
```

```{r cars, echo=FALSE,warning=FALSE}
the_url <- "https://raw.githubusercontent.com/JeffSackmann/tennis_atp/master/atp_matches_2019.csv"
Sackmann_Table <- read.csv(the_url, quote = "", 
                row.names = NULL, 
                stringsAsFactors = FALSE)

#Subet only Grand Slam matches
ATP_matches_2019 <- Sackmann_Table %>% filter(tourney_level == "G")

# Get list of all player Names
# --> Should it be player IDs instead???
#unique(ATP_matches_2019$winner_id)
ATP_2019_GS_Names <- union(ATP_matches_2019$winner_name, ATP_matches_2019$loser_name)

bubble_list <- vector(mode="list", length=length(ATP_2019_GS_Names))
names(bubble_list) <- ATP_2019_GS_Names

###############################################################################
# Code for proportions

for(tag in 1:length(ATP_2019_GS_Names)){
  name <- ATP_2019_GS_Names[tag]
  
  dat1 <- ATP_matches_2019 %>%
    filter(winner_name == name) %>%
    mutate(Service_Games_Won = w_SvGms - w_bpFaced + w_bpSaved,
           Break_points_Won = l_bpFaced - l_bpSaved) %>%
    summarise(n = n(), T_Service_Games_Won = sum(Service_Games_Won, na.rm = T),
              T_Service_Games = sum(w_SvGms, na.rm = T),
              T_Break_points_Won = sum(Break_points_Won, na.rm = T),
              T_opp_Service_Games = sum(l_SvGms, na.rm = T))
  
  dat2 <- ATP_matches_2019 %>%
    filter(loser_name == name) %>%
    mutate(Service_Games_Won = l_SvGms - l_bpFaced + l_bpSaved,
           Break_points_Won = w_bpFaced - w_bpSaved) %>%
    summarise(n = n(), T_Service_Games_Won = sum(Service_Games_Won, na.rm = T),
              T_Service_Games = sum(l_SvGms, na.rm = T),
              T_Break_points_Won = sum(Break_points_Won, na.rm = T),
              T_opp_Service_Games = sum(w_SvGms, na.rm = T))
  
  dat <- rbind(dat1, dat2)
  bubble_list[[tag]] <- dat
  
}

to_plot_dat <- data.frame(matrix(nrow = length(ATP_2019_GS_Names), ncol = 4))
to_plot_dat[,1] <- ATP_2019_GS_Names


for (index in 1:length(bubble_list)){
  player_dat <- bubble_list[[index]] %>%
    summarise_each(sum) %>%
    mutate(Prop_Service_Games_Won = T_Service_Games_Won/T_Service_Games ,
           Prop_Break_points_Won = T_Break_points_Won/T_opp_Service_Games)
  
  to_plot_dat[index,2] <- player_dat$n
  to_plot_dat[index,3] <- player_dat$Prop_Service_Games_Won
  to_plot_dat[index,4] <- player_dat$Prop_Break_points_Won
}

colnames(to_plot_dat) <- c("Player_Name", "Matches", "SvGmsW_Prop", "BPW_Prop")

get.last.name <- function(name) {
  d <- lapply(ifelse(grepl(",",name),strsplit(name,","),strsplit(name," ")),`[[`,2)
  lapply(strsplit(gsub("^ ","",d), " "),`[[`,1)
}

to_plot_dat <- to_plot_dat %>%
  left_join(
    ATP_matches_2019 %>%
      group_by(winner_name) %>%
      dplyr::summarise(Matches_Won=n()),
   by = c('Player_Name' = 'winner_name') 
  ) %>% 
  mutate(Matches_Won = ifelse(is.na(Matches_Won), 0, Matches_Won),
         SvGmsW_Prop = as.numeric(SvGmsW_Prop),
         BPW_Prop = as.numeric(BPW_Prop),
         Last_Name = get.last.name(Player_Name)
         ) %>%
  arrange(desc(Matches))

```

After some data cleaning and aggregating steps on match summary data found on Jeff Sackmann's [Github page](https://github.com/JeffSackmann/tennis_atp/blob/master/README.md), we produce the following table that combines the results from all 4 grand slam tournaments from the 2019 season.


```{r table, echo=FALSE}
A <- to_plot_dat %>%
  arrange(desc(Matches_Won)) %>%
  select(Player_Name, Matches_Won, SvGmsW_Prop, BPW_Prop)

knitr::kable(A[1:10,], caption = "Grand Slam data")
```

During the grand slams Nadal, Djokovic and Federer are ranked 1, 2 and 3 respectively in terms of matches won. The other two columns we calculated represents the proportion of service games defended (**SvGmsW_Prop**), and the proportion of opponent serves broken (**BPW_Prop**).


### Comparing the big 3 to other ATP players via Bubble plots
Using the two metrics (proportion of service games defended and the proportion of opponent serves broken) we mapped out the defensive and offensive prowess of the Big 3 compared to the abilities of the other ATP players in the graphic below. We placed these two metrics on the x- and y-axis respectively. Additionally, to represent the success of the ATP players, we added an extra dimension that represents number of matches won by each player. The larger the *bubble* is, the more grand slam matches a player has won.

```{r pressure2,echo=FALSE,  warning=FALSE}

p <- ggplot(to_plot_dat[-c(1,48,64),], aes(x=SvGmsW_Prop*100, y=BPW_Prop*100, size = Matches_Won, label = Player_Name)) + 
  geom_point(alpha = 0.8, color = "#008080") +
  
  #Add labels for Djoker, Roger and Rafa
  #geom_text_repel(aes(label=ifelse(SvGmsW_Prop > 0.886486 & BPW_Prop > 0.2635869 , 
  #                                 as.character(Last_Name),'')
  #), size = 4.0, nudge_y = -4) +
  
  geom_text(data =  to_plot_dat %>% filter(Last_Name == "Djokovic"), 
            aes(label=as.character(Last_Name)), size = 4, color = "blue",
            nudge_x = -4.5) +
  
  geom_text(data =  to_plot_dat %>% filter(Last_Name == "Nadal"), 
            aes(label=as.character(Last_Name)), size = 4, color = "red",
            nudge_y = -3) +
  geom_text(data =  to_plot_dat %>% filter(Last_Name == "Federer"), 
            aes(label=as.character(Last_Name)), size = 4, color = "forestgreen",
            nudge_y = 3) +
  
  #Adjust size of bubble points and legend title
  scale_size(range = c(.1, 8), name="Matches Won") +
  
  #Colour Nadal red
  geom_point(data=to_plot_dat[1, ], aes(x=SvGmsW_Prop*100, y=BPW_Prop*100, size = Matches_Won),
             colour="red", alpha = 0.7) +
  
  #Colour Federer in Green
  geom_point(data=to_plot_dat[3, ], aes(x=SvGmsW_Prop*100, y=BPW_Prop*100, size = Matches_Won),
             colour="forestgreen", alpha = 0.7) +
  
  #Colour Djoker in Blue
  geom_point(data=to_plot_dat[2, ], aes(x=SvGmsW_Prop*100, y=BPW_Prop*100, size = Matches_Won),
             colour="blue", alpha = 0.7) +
  
  #Add titles
  ggtitle("2019 Grand Slam Season") + 
  xlab("% Service Games Held") + ylab("% Service Games Broken") +
  theme_classic() +
  
  theme(panel.background = element_rect(fill = "#F1F8F7", # background colour
                                        colour = "black", # border colour
                                        size = 0.5, linetype = "solid"),
        plot.title=element_text(size = rel(1.6),
                                face = "bold", hjust = 0.5),
        legend.position = "bottom",
        legend.background = element_rect(colour = "gray"),
        legend.key = element_rect(fill = "#F1F8F7", linetype = 1),
        axis.title = element_text(face = "bold", size = 13)) +
  
  guides(size = guide_legend(override.aes = list(colour = list("#008080"))))

p


```

From the above bubble plot, we make some very intuitive remarks. 

1. Larger bubbles are found on the top right corner of the plot. This indicates that players who are good at both defending their serve and at breaking their opponent serve tend to win more tennis matches. Unsurprisingly, since the Big 3 are good at both, their bubbles are found in the top right corner. 

2. ATP players tend to be better at holding serve than they are at breaking their opponent's serve. Many of the top players hold serve at a rate of 80%, while most players are unable to break their opponent at a rate of 25%.

3. ATP players that are very good at holding serve, but are average at breaking serve still tend to win a sizeable amount of matches. Conversely, players that do not defend their serve well, but who excel at breaking their opponent tend to not win many matches.




For those who like interactive plots as much as I do, I've produced the same plot using the R package ``gplotly``.

```{r pressure,echo=FALSE,  warning=FALSE}

p <- ggplot(to_plot_dat[-c(1,2,3),], aes(x=SvGmsW_Prop*100, y=BPW_Prop*100, size = Matches_Won, label = Player_Name)) + 
  geom_point(alpha = 0.8, color = "#008080") +
  
  #Add labels for Djoker, Roger and Rafa
  #geom_text_repel(aes(label=ifelse(SvGmsW_Prop > 0.886486 & BPW_Prop > 0.2635869 , 
  #                                 as.character(Last_Name),'')
  #), size = 4.0, nudge_y = -4) +
  
  geom_text(data =  to_plot_dat %>% filter(Last_Name == "Djokovic"), 
            aes(label=as.character(Last_Name)), size = 4, color = "blue",
            nudge_x = -6.5) +
  
  geom_text(data =  to_plot_dat %>% filter(Last_Name == "Nadal"), 
            aes(label=as.character(Last_Name)), size = 4, color = "red",
            nudge_y = -4) +
  geom_text(data =  to_plot_dat %>% filter(Last_Name == "Federer"), 
            aes(label=as.character(Last_Name)), size = 4, color = "forestgreen",
            nudge_y = 3) +
  
  #Adjust size of bubble points and legend title
  scale_size(range = c(.1, 8), name="Matches Played") +
  
  #Colour Djokovic blue
  geom_point(data=to_plot_dat[2, ], aes(x=SvGmsW_Prop*100, y=BPW_Prop*100, size = Matches_Won),
             colour="blue", alpha = 0.7) +
  
  #Colour Federer in Green
  geom_point(data=to_plot_dat[3, ], aes(x=SvGmsW_Prop*100, y=BPW_Prop*100, size = Matches_Won),
             colour="forestgreen", alpha = 0.7) +
  
  #Colour Rafa in Red
  geom_point(data=to_plot_dat[1, ], aes(x=SvGmsW_Prop*100, y=BPW_Prop*100, size = Matches_Won),
             colour="red", alpha = 0.7) +
  
  #Add titles
  ggtitle("2019 Grand Slam Season") + 
  xlab("% Service Games Held") + ylab("% Service Games Broken") +
  theme_classic() +
  
  theme(panel.background = element_rect(fill = "#F1F8F7", # background colour
                                        colour = "black", # border colour
                                        size = 0.5, linetype = "solid"),
        plot.title=element_text(size = rel(1.6),
                                face = "bold", hjust = 0.5),
        legend.position = "bottom",
        legend.background = element_rect(colour = "gray"),
        legend.key = element_rect(fill = "#F1F8F7", linetype = 1),
        axis.title = element_text(face = "bold", size = 13)) +
  
  guides(size = guide_legend(override.aes = list(colour = list("#008080"))))

ggplotly(p, tooltip = c("size", "Player_Name"))
```

All R codes to produce this analysis can be found on my [Github page](https://github.com/petertea96/Tennis-Data-Analysis/blob/master/Bubble%20chart.R).
