---
title: "Economic Time-series Data"
author: "Samar"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


#Loading packages


```{r,warning=FALSE,message=FALSE}
library(tidyverse)
library(tidyquant)
library(scales)
```

#Get Data
```{r,warning=FALSE,message=FALSE}
fred_raw <- tq_get(c("RSXFSN",  # Advance retail sales
                     "GDPC1",  # GDP
                     "ICSA",  # Initial unemployment claims
                     "FPCPITOTLZGUSA",  # Inflation
                     "UNRATE",  # Unemployment rate
                     "USREC",   # Recessions
                     "BOPGSTB",#Trade Balance: Goods and Services,Balance of Payments Basis
                     "BOPTEXP",#Export of Goods and Services,Balance of Payment Basis
                     "BOPTIMP",#Import of Goods and Services,Balance of Payment Basis
                     "UMCSENT",#Consumer Sentiment
                     "PERMIT",#Housing Starts and Building Permits
                     "PI",#Personal Income
                     "PCE",#Personal Consumption Expenditures
                     "DGORDER",#Manufacturers' New Orders: Durable Goods
                     "T10Y2Y",#10-Year Treasury Constant Maturity Minus 2-Year Treasury Constant Maturity
                     "TCU",#Capacity Utilization: Total Industry
                     "INDPRO"),#Industrial Production Index
                   get = "economic.data",  # Use FRED
                   from = "1990-01-01")
```
#Recession

```{r,warning=FALSE,message=FALSE}
recessions_tidy <- fred_raw %>% 
  filter(symbol == "USREC") %>% 
  mutate(recession_change = price - lag(price))
recessions_tidy

recessions_start_end <- fred_raw %>% 
  filter(symbol == "USREC") %>% 
  mutate(recession_change = price - lag(price)) %>% 
  filter(recession_change != 0)
recessions_start_end

recessions <- tibble(start = filter(recessions_start_end, recession_change == 1)$date,
                     end = filter(recessions_start_end, recession_change == -1)$date)
recessions




```
#Retail Sales
```{r,warning=FALSE,message=FALSE}
options(scipen=999)
retail_sales <- fred_raw %>% 
  filter(symbol == "RSXFSN")

retail_sales

ggplot(retail_sales,aes(x=date,y=price))+
geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#FF4136", size = 0.5) +
  annotate(geom = "label", x = as.Date("2010-01-01"), y = 10000, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 500000, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
  scale_y_continuous(labels = comma) +
  labs(y = "Retail Sales",
       x = NULL,
       title = "Retail Sales",
       subtitle = "Monthly data",
       caption = "Source: US Employment and Training Administration and FRED") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```
# Unemployment Claims 

```{r,warning=FALSE,message=FALSE}
options(scipen=999)
unemployment_claims_only <- fred_raw %>% 
  filter(symbol == "ICSA")

unemployment_claims_only


ggplot(unemployment_claims_only, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#FF4136", size = 0.5) +
  annotate(geom = "label", x = as.Date("2010-01-01"), y = 1000000, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 7000000, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
  scale_y_continuous(labels = comma) +
  labs(y = "Initial unemployment claims",
       x = NULL,
       title = "Initial unemployment claims",
       subtitle = "Weekly data",
       caption = "Source: US Employment and Training Administration and FRED") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```
#GDP

```{r,warning=FALSE,message=FALSE}
options(scipen=999)
gdp_only <- fred_raw %>% 
  filter(symbol == "GDPC1")

ggplot(gdp_only, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#0074D9", size = 1) +
  
  annotate(geom = "label", x = as.Date("2010-01-01"), y = 2000, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 20000, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
  
  scale_y_continuous(labels = dollar) +
  labs(y = "Billions of 2012 dollars",
       x = NULL,
       title = "US Gross Domestic Product",
       subtitle = "Quarterly data; real 2012 dollars",
       caption = "Source: U.S. Bureau of Economic Analysis") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```
#Unemployment Rate

```{r,warning=FALSE,message=FALSE}
options(scipen=999)
Unemployment_Rate <- fred_raw %>% 
  filter(symbol == "UNRATE")

Unemployment_Rate


ggplot(Unemployment_Rate, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#FF4136", size = 0.5) +
  annotate(geom = "label", x = as.Date("2010-01-01"), y =0, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 20, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
  scale_y_continuous(labels = comma) +
  labs(y = "Unemployment_Rate",
       x = NULL,
       title = "Unemployment_Rate",
       subtitle = "Monthly data",
       caption = "Source: U.S. Bureau of Labor Statistics") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```
#Inflation

```{r,warning=FALSE,message=FALSE}
options(scipen=999)
Inflation <- fred_raw %>% 
  filter(symbol == "FPCPITOTLZGUSA")

Inflation


ggplot(Inflation, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#FF4136", size = 0.5) +
  annotate(geom = "label", x = as.Date("2010-01-01"), y =-2.5, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 15, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
  scale_y_continuous(labels = comma) +
  labs(y = "Percent",
       x = NULL,
       title = "Inflation",
       subtitle = "Annual data",
       caption = "Source: World Bank") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```
#International Trade in Goods and Services: Trade Balance

```{r,warning=FALSE,message=FALSE}
options(scipen=999)
Trade_Balance_Good_Services_Payment_Basis <- fred_raw %>% 
  filter(symbol == "BOPGSTB")

Trade_Balance_Good_Services_Payment_Basis


ggplot(Trade_Balance_Good_Services_Payment_Basis, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  annotate(geom = "label", x = as.Date("2010-01-01"), y =-70000, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 0, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
  
  geom_line(color = "#FF4136", size = 0.5) +
 
  scale_y_continuous(labels = comma) +
  labs(y = "Millions of Dollars",
       x = NULL,
       title = "Trade_Balance:Good and Services,Balance of Payments Basis",
       subtitle = "Monthly data",
       caption = "Source: U.S. Bureau of Economic Analysis") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```
#Exports of Goods and Services

```{r,warning=FALSE,message=FALSE}
options(scipen=999)
Export_Balance_Good_Services_Payment_Basis <- fred_raw %>% 
  filter(symbol == "BOPTEXP")

Export_Balance_Good_Services_Payment_Basis


ggplot(Export_Balance_Good_Services_Payment_Basis, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#FF4136", size = 0.5) +
  
  annotate(geom = "label", x = as.Date("2010-01-01"), y =0, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 200000, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
 
  scale_y_continuous(labels = comma) +
  labs(y = "Millions of Dollars",
       x = NULL,
       title = "Export of Good and Services,Balance of Payments Basis",
       subtitle = "Monthly data",
       caption = "Source: U.S. Bureau of Economic Analysis") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```
#Import of Goods and Services

```{r,warning=FALSE,message=FALSE}
options(scipen=999)
Import_Balance_Good_Services_Payment_Basis <- fred_raw %>% 
  filter(symbol == "BOPTIMP")

Import_Balance_Good_Services_Payment_Basis


ggplot(Import_Balance_Good_Services_Payment_Basis, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#FF4136", size = 0.5) +
  
  annotate(geom = "label", x = as.Date("2010-01-01"), y =0, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 300000, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
 
  scale_y_continuous(labels = comma) +
  labs(y = "Millions of Dollars",
       x = NULL,
       title = "Import of Good and Services,Balance of Payments Basis",
       subtitle = "Monthly data",
       caption = "Source: U.S. Bureau of Economic Analysis") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```
#Consumer Sentiment

```{r,warning=FALSE,message=FALSE}
options(scipen=999)
Consumer_Sentiment <- fred_raw %>% 
  filter(symbol == "UMCSENT")

Consumer_Sentiment


ggplot(Consumer_Sentiment, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#FF4136", size = 0.5) +
  
  annotate(geom = "label", x = as.Date("2010-01-01"), y =0, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 120, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
 
  scale_y_continuous(labels = comma) +
  labs(y = "Index 1966:Q1=100",
       x = NULL,
       title = "Consumer_Sentiment",
       subtitle = "Monthly data",
       caption = "Source: University of Michigan") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```

#Housing Starts and Building Permits
```{r,warning=FALSE,message=FALSE}
options(scipen=999)
PERMIT <- fred_raw %>% 
  filter(symbol == "PERMIT")

PERMIT


ggplot(PERMIT, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#FF4136", size = 0.5) +
  
  annotate(geom = "label", x = as.Date("2010-01-01"), y =0, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 120, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
 
  scale_y_continuous(labels = comma) +
  labs(y = "Thousands of Units",
       x = NULL,
       title = "New Private Housing Units Authorized by Building Permits",
       subtitle = "Monthly data",
       caption = "Source: U.S. Census Bureau") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```
#Personal Income

```{r,warning=FALSE,message=FALSE}
options(scipen=999)
PI <- fred_raw %>% 
  filter(symbol == "PI")

PI


ggplot(PI, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#FF4136", size = 0.5) +
  
  annotate(geom = "label", x = as.Date("2010-01-01"), y =0, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 120, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
 
  scale_y_continuous(labels = comma) +
  labs(y = "Billions of Dollars",
       x = NULL,
       title = "Personal Income",
       subtitle = "Monthly data",
       caption = "Source: U.S. Bureau of Economic Analysis") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```
#Personal Consumption Expenditures

```{r,warning=FALSE,message=FALSE}
options(scipen=999)
PCE <- fred_raw %>% 
  filter(symbol == "PCE")

PCE

ggplot(PCE, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#FF4136", size = 0.5) +
  
  annotate(geom = "label", x = as.Date("2010-01-01"), y =0, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 16000, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
 
  scale_y_continuous(labels = comma) +
  labs(y = "Billions of Dollars",
       x = NULL,
       title = "Personal Consumption Expenditures",
       subtitle = "Monthly data",
       caption = "Source: U.S. Bureau of Economic Analysis") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```
#Durable Goods Order

```{r,warning=FALSE,message=FALSE}
options(scipen=999)
DGORDER <- fred_raw %>% 
  filter(symbol == "DGORDER")

DGORDER

ggplot(DGORDER, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#FF4136", size = 0.5) +
  
  annotate(geom = "label", x = as.Date("2010-01-01"), y =100000, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 300000, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
 
  scale_y_continuous(labels = comma) +
  labs(y = "Billions of Dollars",
       x = NULL,
       title = "Manufacturers' New Orders: Durable Goods ",
       subtitle = "Monthly data",
       caption = "Source: U.S. Census Bureau") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```
#Yield:10-Year Treasury Constant Maturity Minus 2-Year Treasury Constant Maturity

```{r,warning=FALSE,message=FALSE}
options(scipen=999)
T10Y2Y <- fred_raw %>% 
  filter(symbol == "T10Y2Y")

T10Y2Y

ggplot(T10Y2Y, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#FF4136", size = 0.5) +
  
  annotate(geom = "label", x = as.Date("2010-01-01"), y =-0.2, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 1.8, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
 
  scale_y_continuous(labels = comma) +
  labs(y = "Percent",
       x = NULL,
       title = "10-Year Treasury Constant Maturity Minus 2-Year Treasury Constant Maturity",
       subtitle = "Monthly data",
       caption = "Source: U.S. Census Bureau") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```
# Capacity Utilization: Total Industry

```{r,warning=FALSE,message=FALSE}
options(scipen=999)
TCU <- fred_raw %>% 
  filter(symbol == "TCU")

TCU

ggplot(TCU, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#FF4136", size = 0.5) +
  
  annotate(geom = "label", x = as.Date("2010-01-01"), y =60, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 80, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
 
  scale_y_continuous(labels = comma) +
  labs(y = "Percent of Capacity",
       x = NULL,
       title = "Capacity Utilization: Total Industry",
       subtitle = "Monthly data",
       caption = "Source: Board of Governors of the Federal Reserve System (US)") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))
```
# Industrial Production

```{r,warning=FALSE,message=FALSE}
options(scipen=999)
INDPRO <- fred_raw %>% 
  filter(symbol == "INDPRO")

INDPRO

ggplot(INDPRO, aes(x = date, y = price)) +
  geom_rect(data = recessions, 
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE, fill = "#B10DC9", alpha = 0.3) +
  geom_line(color = "#FF4136", size = 0.5) +
  
  annotate(geom = "label", x = as.Date("2010-01-01"), y =0, 
           label = "The Great Recession", size = 3, family = "Roboto Condensed") +
  annotate(geom = "label", x = as.Date("2020-01-01"), y = 120, 
           label = "COVID-19", size = 3, family = "Roboto Condensed", hjust = 1) +
 
  scale_y_continuous(labels = comma) +
  labs(y = "Index 2012=100",
       x = NULL,
       title = "Industrial Production Index",
       subtitle = "Monthly data",
       caption = "Source: Board of Governors of the Federal Reserve System (US)") +
  theme_bw(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(face = "bold"))



```



